---
title: "Final Project R Notebook"
output: html_document
date: "2026-05-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r cars}
# load all the necessary packages ( not just for this assignment)
library(knitr)
library(readr)
library(dplyr)
library(tidyr)
library(tidyverse)
library(lmtest)
library(stargazer)
library(sandwich)
library(car)
library(tsDyn)
library(vars)

```


```{r}

# Download the csv file
time_series_data = read.csv('/Users/khan/time_series_data.csv')
  
```

First regression:

```{r}

# Overall/General/Mean Sentiment Score Regression:
its_general <- lm(genscore ~ post_911 + sent + unrate +
                  comedypop + crimepop + fanpop + horpop + warpop + scipop + thrillpop, 
                  # all genre box office proportions
                  data = time_series_data) # using the time_series dataset
summary(its_general) # summary

```
Second regression

```{r}

# WAR sentiment score regression
its_war <- lm(warscore ~ post_911 + unrate + sent + comedypop + warpop + crimepop +
                fanpop + horpop + warpop + scipop + thrillpop,
                 data = time_series_data)
summary(its_war) # summarize

```
Third regression

For the genre-interaction model, we are going to use the "tidyr" pivot_longer function to transform our time series dataset, allowing us to compare 9/11 impact across different genres — Romcom, Fantasy, Horror, Crime-Thriller, and War.
It is kind of a pivot table, already transformed into a dataframe.

```{r}

# Pivoting to long format
genre_data <- pivot_longer(time_series_data, cols = c(comscore, fanscore, horscore, crimescore, warscore), 
                           # listing our genre-specific scores
                           names_to = "genre", # a separate column "genre" appears, 
                           values_to = "score") # assigning the "genre_scores" to each "genre"
# 'score' becomes our dependent variable

# The regression acquires genre interactions with post_911, comedy is the reference category

its_genre <- lm(score ~ post_911*genre + unrate + sent
                + comedypop + crimepop + fanpop + horpop + warpop + scipop + thrillpop,
                data = genre_data)
summary(its_genre) # summarize

```

Results Interpretation: After 9/11, crime thrillers saw a decline -- in comparison to romcom movies -- with below 1% significance; their review sentiment score went 0.11 points lower than that of romcoms, holding the genre differences, box office genre dominances, unemployment rate, and UMich consumer sentiment constant. Interestingly, the post 911 dummy -- belonging to the romcom genre as the reference category -- seemed to rise after 9/11. Its p-value is just below 10%, so we'll conclude this occurred by chance. So, for crime thrillers, the politically-charged category of films, post-9/11 overall decline was −0.070 (=0.037−0.107). Finally, war films' sentiment scores did not exhibit a significant value, meaning 9/11 did not have a significant impact on it.


```{r}

# checking what the stargazer variables will look like not to mess anything up
stargazer(its_general, its_war, its_genre, type = "text") 
stargazer(its_general, its_war, its_genre, # what models we want to display
          title = "The Impact of 9/11 on Film Review Sentiment", # the title
          column.labels = c("General Model", "War Model", "Genre Interactions"), # each model's name
          dep.var.labels = c("General Sentiment Score", "War Sentiment Score", "Genre Sentiment Score"),
          dep.var.caption = "", # excludes the name 'Dependent Variables'
          covariate.labels = c("Post 9/11", # post_911 dummy variable
                             "Crime Thriller Score", # genrecrimescore 
                             "Fantasy/Sci-Fi Score", # genrefanscore
                             "Horror Score",  # genrehorscore
                             "War Score", # genrewarscore,
                             "Unemployment Rate", # unrate,
                             "UMich Consumer Sentiment", # sent
                             "Comedy Dominance", # comedypop
                             "Crime Dominance", # crimepop
                             "Fantasy Dominance", # fanpop
                             "Horror Dominance", # horpop
                             "War Dominance", # warpop
                             "Sci-Fi Dominance", # scipop
                             "Thriller Dominance", # thrillpop
                             "Post 9/11 * Crime Thriller", # post_911:genrecrimscore
                             "Post 9/11 * Fantasy/Sci-Fi", # post_911:genrefanscore
                             "Post 9/11 * Horror", # post_911:genrehorscore 
                             "Post 9/11 * War", # post_911:genrewarscore 
                             "Constant" # intercept
                             ),
          type = "html", # type of output
          out = "Three Regressions.html") # name of the file

```

```{r}

# I wanted to make the columns wider, so I asked AI
html_content <- readLines("Three Regressions.html") 
css <- "
<style>
  table {
    width: 50%;
    margin: auto;
  }
</style>
"
html_content <- c(css, html_content)
writeLines(html_content, "Three Regressions.html")

```

